Bricking up and overgrowing backdoors in Intrusion Detection Systems

### 2019-09-08 ex2/paper/paper.tex
Small corrections in my part of paper

As shown in \autoref{fig:pdp_ttl}, the \gls{mui} for both \gls{rf} and \gls{dl} shows a deep notch for certain low values of stdev(TTL). As discussed above, normal traffic is very unlikely to have deviating \gls{ttl} values for different packets. One would therefore expect this feature to have a negligible influence on the classification result.In our case, this notch indicates the existance of the backdoor.

We have to note, however, that inconsistent behaviour of the \gls{mui} detected using \gls{pdp} or \gls{ale} plots does not necessarily result from poisoning activity.  For example, \autoref{fig:ttlmean} shows the mean(TTL) feature in forward direction. The models show a clear dependence of the mean \gls{ttl} value of incoming packets. In our case, this behaviour results from the non-equal distribution of \gls{ttl} values of attack and non-attack traffic in both the UNSW-NB15 and CIC-IDS-2017 datasets.

Independent of the origin of such patterns, they might be exploited for masquerading attacks and thus are clearly unwanted. The use of \gls{pdp} and \gls{ale} plots therefore provides a convenient possibility for analyzing \gls{ml} models.
\subsection{Pruning Techniques}
\subsubsection{Neural Network Pruning}

@@

\subsection{Neural Network Fine-Tuning}
Next to pruning a frequently used technique to remove backdoors from a trained model is fine-tuning (\todo{ref}). For fine-tuning training of the \gls{mui} is continued with the validation set, hence reinforcing correct decisions and causing the \gls{mui} to gradually forget backdoors.

@@

Indeed, in (\todo{ref}) it is argued that since fine-tuning removes backdoors from neurons that are activated by the validation set, fine-tuning is the ideal complement for pruning. They thus propose fine-pruning as a combination of pruning and fine-tuning.

\autoref{fig:finetuning} show the backdoor efficacy when continuing training of our trained model for CIC-IDS-2015 with the validation set, hence without backdoored samples. 
\autoref{fig:finetuning} indicates clearly that, unfortunately, similar to pruning fine-tuning is uneffective for cleaning the \gls{mui} from backdoors in our case.

Note that the use of fine-tuning exclusively makes  sense if both the computational effort and the required training set size for fine-tuning are substantially lower than for the original training procedure. Hence, a a decay of backdoor efficacy as depicted in \autoref{fig:finetuning} cannot be considered sufficient.

>>>>

As shown in \autoref{fig:pdp_ttl}, the \glspl{pdp} for both \gls{rf} and \gls{dl} show a deep notch for certain low values of stdev(TTL). As discussed above, normal traffic is very unlikely to have deviating \gls{ttl} values for different packets. In contrast to \autoref{fig:pdp_ttl}, one would therefore expect this feature to have a negligible influence on the classification result. In our case, this notch thus points to the existance of the backdoor we embedded.

It has to be noted, however, that inconsistent behaviour of the \gls{mui}, detected using \gls{pdp} or \gls{ale} plots, does not necessarily result from poisoning activity.  For example, \autoref{fig:ttlmean} shows the mean(TTL) feature in forward direction. The models show a clear dependence of the mean \gls{ttl} value of incoming packets, which is similarly counter-intuitive as for the feature discussed above. In our case, this behaviour results from the non-equal distribution of \gls{ttl} values of attack and non-attack traffic in both the UNSW-NB15 and CIC-IDS-2017 datasets.

Independent of the origin of such patterns, they might be exploited for masquerading attacks and thus are clearly unwanted. The use of \gls{pdp} and \gls{ale} plots therefore provides a convenient possibility for analyzing \gls{ml} models for such patterns.
\subsection{Pruning Techniques}
\subsubsection{Neural Network Pruning}

@@

\subsection{Neural Network Fine-Tuning}
Besides pruning, a frequently used technique for removing backdoors from a trained model is fine-tuning (\todo{ref}). For fine-tuning, training of the \gls{mui} is continued with the validation set, hence reinforcing correct decisions and causing the \gls{mui} to gradually forget backdoors.

@@

Indeed, in (\todo{ref}) it is argued that since fine-tuning removes backdoors from neurons that are activated by the validation set, fine-tuning is the ideal complement for pruning, which removes backdoors from neurons that are not activated by the validation set. They thus propose fine-pruning as a combination of pruning and fine-tuning.

\autoref{fig:finetuning} shows the backdoor efficacy when continuing training of our trained model for CIC-IDS-2015 with the validation set, hence without backdoored samples. 
\autoref{fig:finetuning} indicates clearly that, unfortunately, similar to pruning fine-tuning is uneffective for cleaning the \gls{mui} from backdoors in our case.

The use of fine-tuning exclusively makes  sense if both the computational effort and the required training set size for fine-tuning are substantially lower than for the original training procedure. Hence, a decay of the backdoor efficacy as depicted in \autoref{fig:finetuning} cannot be considered sufficient.
